Basic Model formulation

### 2015-02-20 documentation/model_formulation.Rmd
try to visualise attractor of pair-approximation model

>>>>

### Attractor

Visualising the attractor of the system is not as straightforward as for the models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria. In the spatially-explict context the attractor can be visualised if we assume that clustering in the initial state is random, i.e. $q_{11} = \rho_{1}$.

parms$rho_ini <- seq(0, 1, 0.1)
parms$m_ini <- seq(0, 0.2, length = 11)

rho_one_one_ini <- function(prms) {
  out <- sapply(1:nrow(prms), function(x) {
  i <- with(prms[x,], v * L * p * rho_ini^(q-1) )
  ii <- with(prms[x,], ( v*m*h*rho_ini - v*h*m_ini + p*a*L - v*L )*rho_ini^q )
  iii <- with(prms[x,], -m_ini + m*rho_ini + a*h*m*rho_ini^(2+q) - a*h*m_ini*rho_ini^(1+q) - a*L*rho_ini^(1+q) )

  solutions <- data.frame( 
    s1 = (- ii + sqrt( ii^2 - 4*i*iii ) ) / 2 * i,
    s2 = (- ii - sqrt( ii^2 - 4*i*iii ) ) / 2 * i)
  return(solutions)
  })
  return(as.data.frame(t(out)))
}

iterations$rho_11_ini <- unlist(rho_one_one_ini(iterations)$s1)

ini_rho <- function(rho_1, rho_11 = NULL) {
    if(is.null(rho_11)) out <- {
      c(
    rho_1 = rho_1,
    rho_11 = rho_1*rho_1,
    rho_10 = rho_1*(1-rho_1)*2,
    rho_00 = (1-rho_1)*(1-rho_1),
    rho_0 = 1-rho_1
    ) 
    } else {
      out <- c(
    rho_1 = rho_1,
    rho_11 = rho_11,
    rho_10 = rho_1-rho_11,
    rho_00 = 1-2*rho_1+rho_11,
    rho_0 = 1-rho_1
    )
    if(any(out < 0)) out <- c(
    rho_1 = NA,
    rho_11 = NA,
    rho_10 = NA,
    rho_00 = NA,
    rho_0 = NA
    )
    }

  return(out)
}

starting <- sapply(1:nrow(iterations), function(x)  ini_rho(iterations[x, "rho_ini"], iterations[x, "rho_11_ini"])  )

defplot(rho, C(rho, 1), ylab = "mortality")
lines(rho,C(rho, 0.75))
lines(rho,C(rho, 0.5))
lines(rho,C(rho, 0.25))
lines(rho,C(rho, 0))

lines(rho, G(rho, 1), col = "green3")
lines(rho,G(rho, 0.75), col = "green3")
lines(rho,G(rho, 0.5), col = "green3")
lines(rho,G(rho, 0.25), col = "green3")
lines(rho,G(rho, 0), col = "green3")

points(iterations$rho_ini, iterations$m_ini, pch = 20, col = is.na(starting[1,])+1)

  model_parms <- as.list(iterations[iteration,])
  rho_starting <- ini_rho(iterations[iteration, "rho_ini"], iterations[iteration, "rho_11_ini"])
  if(!any(is.na(rho_starting)) ) {

  # running the ode-solver
  runmodel <- ode(y = rho_starting, func = odesys_spex, times = c(1,2), parms = model_parms)

  out <-as.data.frame(runmodel)[dim(runmodel)[1],]
  }  else { 
    out <- c(time = 100,rho_starting)  
  }
  # transfer into ouput and calculate missing rho values
  return(out)
  } -> output

output$rho_11_ini <- iterations$rho_11_ini

G <- function(rho_1, q_01 = "auto", prms = parms) {
  if(q_01 == "auto") q_01 <- 1
  with(prms, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))

C <- function(rho_1, q_01 = 1, prms = parms) with(prms, (m * rho_1) +( (a+v*q_01  ) *rho_1^(1+q)*L*(1-p*q_01))/(1+(a+v*q_01)*h*(rho_1^(1+q)) ) )

arrows(output$rho_ini, C(output$rho_ini, output$rho_11_ini),output$rho_1, C(output$rho_1, output$rho_11), length = 0.05)
